### 2022-11-15 NSF2022/Description/specificaim1v2.tex

\begin{wrapfigure}[11]{r}{0.5\textwidth}
  \vspace{-8pt}
  \centerline{\includegraphics[width=0.5\textwidth]{figures/SA1Figures/AmphiphilicAssembly.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:amphiphilic_assembly} \footnotesize
  Self-organization under an amphiphilic boundary condition. (a)
  Granules first match hydrophobic tails (red), (b) then align
  length-wise. (c) The same pattern of self-organization holds in three
  dimensions, as simulated by Kohl, Corona, and
  Veerapaneni~\cite{koh-cor-che-vee2021}.}
\end{wrapfigure}
In the double-well potential case, equation~\eqref{eqn:phase} is a
nonlinear Allen-Cahn equation. This situation models granules immersed
in a two-phase fluid (e.g., granules in an emulsion) where there is an
interesting interplay between the long-range alignment of granules and
the diffusive interface energy of the free boundary between fluid
phases. On unbounded domains like $\Omega(t)$, Allen-Cahn equations can
have multiple solutions~\cite{Alama1997StationaryLS,
Alikakos2008OnAE, Bronsard1993OnTB, Byeon2014SolutionsOH, Byeon2013OnAP,
Alessio2005ENTIRESI, Trumper2007ExistenceOA, Benci2019MultipleSF}.
To have one solution $\phi$ for each $t$, 
we must reformulate~\eqref{eqn:phase} as
the following initial value problem with transport:
\begin{align}
  \label{eqn:phase_transport}
  \phi_t + \uu \cdot \nabla
  = -\beta \frac{\delta E}{\delta \phi}
  = \beta(\epsilon^2 \Delta \phi - W'(\phi)),
  \quad \phi(\xx,0) = \phi_0(\xx),
\end{align}
for initial data $\phi_0$. The modified system enjoys an energy law
similar to~\eqref{eq:energy_law} with an additional dissipation term
involving the square the Euler-Lagrange derivative $\frac{\delta
E}{\delta \phi}$ and dampening factor $\beta > 0$. There are several
works dealing with the characterization of the geometric evolution of
interfaces under the Allen-Cahn and Cahn-Hilliard
equations~\cite{Christlieb2019CompetitionAC, Gavish2011CurvatureDF,
Dai2019WeakSF, Promislow2017ExistenceBA, Dai2015CompetitiveGE,
Promislow2012CriticalPO, Dai2022GeometricEO, Dai2020MinimizersFT,
Dai2013GeometricEO, Promislow2022UndulatedBI, Gera2017CahnHilliardOS}
and well-posedness
results when coupled to flow in static domains~\cite{Jiang2017TwophaseIF,
Liu2012StrongSF, Giorgini2019WellPosednessOA, Wu2022WellposednessOA,
Gal2010AsymptoticBO, Giorgini2020DiffuseIM, Giorgini2019UniquenessAR}.
The proposed research extends these results to problems with moving
domains. A closely related and strongly developing direction in the area
of nematics liquid crystals concerns the interaction of Landau-de Gennes
functionals with colloids, and studies the qualitative structure of
defects around single or multiple colloidal
particles~\cite{doi:10.1098/rsta.2020.0432, Alama2015MinimizersOT,
Alama2021SaturnRD, PhysRevE.96.042702}.
PI Ryham has several works on
membrane vesicles dealing both with Allen-Cahn formulations and
fluid-vesicle interactions~\cite{QiangDu09, RYHAM20112929, RyCoEi12,
Ryham2017OnTV}.

@@

\begin{quotation}
  \noindent
  \textbf{Problem 2.}
  Consider a collection of two-dimensional amphiphilic granules along a
  closed curve $\mathcal{C}$. What is the sharp interface limit of the
  energy $E$ as the number of granules goes to infinity and their size
  to zero?
\end{quotation}

Sharp interface models represent a membrane vesicle as an
inextensible, continuous, closed surface $\Sigma$ with elastic energy
\begin{align}
  \label{eq:Canham-Helfrich}
  \int_{\Sigma} k_B(H - k_0)^2\, dS,
\end{align}
where $H$ is the mean curvature, $k_0$ is a spontaneous curvature, and
$k_B$ is a bending modulus. One can also account
for spatially dependent spontaneous curvature
\cite{PhysRevE.79.031926,Lowengrub13,mahapatra_saintillan_rangamani_2020}
and for extensibility by
including in~\eqref{eq:Canham-Helfrich} an additional stretching term
$k_A(\alpha - \alpha_0)^2/\alpha_0^2$ where $\alpha$ and $\alpha_0$ are
the deformed and equilibrium area densities respectively
\cite{chabanon2017}.

>>>>

\begin{wrapfigure}[11]{r}{0.5\textwidth}
  \vspace{-8pt}
  \centerline{\includegraphics[width=0.5\textwidth]{figures/SA1Figures/AmphiphilicAssembly.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:amphiphilic_assembly} \footnotesize
  Self-organization under an amphiphilic boundary condition. (a)
  Granules first match hydrophobic tails (red), (b) then align
  lengthwise. (c) The same pattern of self-organization holds in three
  dimensions, as simulated by Kohl, Corona, and
  Veerapaneni~\cite{koh-cor-che-vee2021}.}
\end{wrapfigure}
In the double-well potential case, equation~\eqref{eqn:phase} is a
nonlinear Allen-Cahn equation. This situation models granules immersed
in a two-phase fluid (e.g., granules in an emulsion) where there is an
interesting interplay between the long-range alignment of granules and
the diffusive interface energy of the free boundary between fluid
phases. On unbounded domains like $\Omega(t)$, Allen-Cahn equations can
have multiple solutions~\cite{Alama1997StationaryLS,
Alikakos2008OnAE, Bronsard1993OnTB, Byeon2014SolutionsOH, Byeon2013OnAP,
Alessio2005ENTIRESI, Trumper2007ExistenceOA, Benci2019MultipleSF}.
To have one solution $\phi$ for each $t$, 
we must reformulate~\eqref{eqn:phase} as
the following initial value problem with transport:
\begin{align}
  \label{eqn:phase_transport}
  \phi_t + \uu \cdot \nabla
  = -\beta \frac{\delta E}{\delta \phi}
  = \beta(\epsilon^2 \Delta \phi - W'(\phi)),
  \quad \phi(\xx,0) = \phi_0(\xx),
\end{align}
for initial data $\phi_0$. The modified system enjoys an energy law
similar to~\eqref{eq:energy_law} with an additional dissipation term
involving the square the Euler-Lagrange derivative $\frac{\delta
E}{\delta \phi}$ and dampening factor $\beta > 0$. There are several
works dealing with the characterization of the geometric evolution of
interfaces under the Allen-Cahn and Cahn-Hilliard
equations~\cite{Christlieb2019CompetitionAC, Gavish2011CurvatureDF,
Dai2019WeakSF, Promislow2017ExistenceBA, Dai2015CompetitiveGE,
Promislow2012CriticalPO, Dai2022GeometricEO, Dai2020MinimizersFT,
Dai2013GeometricEO, Promislow2022UndulatedBI, Gera2017CahnHilliardOS}
and well-posedness
results when coupled to flow in static domains~\cite{Jiang2017TwophaseIF,
Liu2012StrongSF, Giorgini2019WellPosednessOA, Wu2022WellposednessOA,
Gal2010AsymptoticBO, Giorgini2020DiffuseIM, Giorgini2019UniquenessAR}.
The proposed research extends these results to problems with moving
domains. A closely related and strongly developing direction in the area
of nematic liquid crystals concerns the interaction of Landau-de Gennes
functionals with colloids, and studies the qualitative structure of
defects around single or multiple colloidal
particles~\cite{doi:10.1098/rsta.2020.0432, Alama2015MinimizersOT,
Alama2021SaturnRD, PhysRevE.96.042702}.
PI Ryham has several works on
membrane vesicles dealing both with Allen-Cahn formulations and
fluid-vesicle interactions~\cite{QiangDu09, RYHAM20112929, RyCoEi12,
Ryham2017OnTV}.

@@

\begin{quotation}
  \noindent
  \textbf{Problem 2.}
  Consider a collection of two-dimensional amphiphilic granules along a
  closed curve $\mathcal{C}$. What is the sharp interface limit of the
  energy $E$ as the number of granules goes to infinity and their size
  to zero?
\end{quotation}
Sharp interface models represent a membrane vesicle as an
inextensible, continuous, closed surface $\Sigma$ with elastic energy
\begin{align}
  \label{eq:Canham-Helfrich}
  \int_{\Sigma} k_B(H - k_0)^2\, dS,
\end{align}
where $H$ is the mean curvature, $k_0$ is a spontaneous curvature, and
$k_B$ is a bending modulus. One can also account
for spatially dependent spontaneous curvature
\cite{PhysRevE.79.031926,Lowengrub13,mahapatra_saintillan_rangamani_2020}
and for extensibility by
including in~\eqref{eq:Canham-Helfrich} an additional stretching term
$k_A(\alpha - \alpha_0)^2/\alpha_0^2$ where $\alpha$ and $\alpha_0$ are
the deformed and equilibrium area densities respectively
\cite{chabanon2017}, and $k_A$ is a stretching modulus.

### 2022-11-15 NSF2022/Description/specificaim2.tex

To briefly describe IE formulations of PDEs, we consider the Stokes
equation~\eqref{eqn:stokes} as an example. Its solution can be written
as a convolution 
\begin{align}
  \uu(\xx) = \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, dS_\yy,
  \label{eqn:conv}
\end{align}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega(t) \rightarrow \RR^2$
is an unknown density function. The integral
representation~\eqref{eqn:conv} is known as a layer potential.
%This is analogous to the problem of
%finding the global electrostatic potential $u$ with surface charge
%density $\sigma$
%\begin{align*}
%  u(x) = -\frac{1}{4\pi}\int_{\partial \Omega(t)} 
%    \frac{1}{|\xx - \yy|}\sigma(y) \, ds_\yy.
%\end{align*}
The IE reformulation has several immediate advantages including a
dimension reduction since $\ssigma$ is defined only on $\partial
\Omega(t)$, and the far field boundary condition of the PDE is
automatically satisfied. The density function $\ssigma$ is chosen to
satisfy the boundary condition, and this results is an IE 
\begin{align}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, dS_\yy,
    \quad \xx \in \partial\Omega(t),
  \label{eqn:bie}
\end{align} where $\gg$ is the boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

@@

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 4.} How can novel quadrature, time adaptivity, fast
  summation, and collision-detection be combined to accurately and
  efficiently solve the governing
  equations~\eqref{eq:RBT}--\eqref{eqn:stressbalance} with dense
  suspensions of two- and three-dimensional granules in bounded and
  unbounded domains?
\end{quotation}

In experiments, it is known that interactions between bounding walls and
granules cannot be ignored~\cite{qua-gan-you2021,
KaouiHartingMisbah2011_PRE, her-sto-gra2005}. Therefore, the necessary
modification to the code will be implemented so that confinement can be
considered. This modification requires introducing an additional layer
potential on the bounding walls, and elements of this modification have
already been incorporated (Figure~\ref{fig:stenosis}). However, to
consider rheological properties of the suspensions, a periodic domain,
such as a Couette device, will be considered. Such multiply-connected
domains require additional modifications to the IE formulation. PI
Quaife is qualified to perform this task since he has simulated
suspensions with confinement for several projects including dynamic
suspensions~\cite{qua-bir2014, bys-sha-qua2020} and porous
media~\cite{chi-moo-qua2020, dea-qua-bir-jua2018, moo-che-chi-qua2022,
qua-gan-you2021}.

@@

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 5.} How can IEs be coupled with a Newton method to
  solve the non-linear PDE~\eqref{eqn:phase}?
\end{quotation}

For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. This requires solving the variable coefficient PDE
$-\epsilon^2 \Delta v + W''(\phi^{N}) v = F[\phi^N]$, which is
challenging with an integral equation since a fundamental solution is
not readily available. However, by choosing a double-well potential with
$W''(\phi_0) = W''(\phi_1) = \beta$, we propose to instead use an
inexact Newton's method where $(F')^{-1}$ is approximated with the
solution of $-\epsilon^2 \Delta v + \beta v = F[\phi^N]$. We test the
efficacy of the exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1) \subset \RR$ by discretizing $F'$
with a second-order finite difference stencil. The double-well potential
is illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the same solution. Figure~\ref{fig:CA}(b)
shows the norm of the residual at each Newton
\begin{wrapfigure}[18]{l}{1.8in}
  \vspace{-7pt}
  \centering
  \input{figures/doublewell.tikz} 
  \vspace{-10pt}
  \caption{\label{fig:CA} \footnotesize (a) A double-well potential. (b)
  The convergence of exact and inexact Newton's methods applied
  to~\eqref{eqn:phaseRoot}.}
\end{wrapfigure}
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

>>>>

To briefly describe IE formulations of PDEs, we consider the Stokes
equation~\eqref{eqn:stokes} as an example. Its solution can be written
as a convolution 
\begin{align}
  \uu(\xx) = \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, dS_\yy,
  \label{eqn:conv}
\end{align}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega(t) \rightarrow \RR^2$
is an unknown density function. The integral
representation~\eqref{eqn:conv} is known as a layer potential.
%This is analogous to the problem of
%finding the global electrostatic potential $u$ with surface charge
%density $\sigma$
%\begin{align*}
%  u(x) = -\frac{1}{4\pi}\int_{\partial \Omega(t)} 
%    \frac{1}{|\xx - \yy|}\sigma(y) \, ds_\yy.
%\end{align*}
The IE reformulation has several immediate advantages including a
dimension reduction since $\ssigma$ is defined only on $\partial
\Omega(t)$, and the far-field boundary condition of the PDE is
automatically satisfied. The density function $\ssigma$ is chosen to
satisfy the boundary condition, and this results is an IE 
\begin{align}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, dS_\yy,
    \quad \xx \in \partial\Omega(t),
  \label{eqn:bie}
\end{align} where $\gg$ is the boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

@@

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 4.} How can novel quadrature, time adaptivity, fast
  summation, and collision-detection be combined to accurately and
  efficiently solve the governing
  equations~\eqref{eq:RBT}--\eqref{eqn:stressbalance} with dense
  suspensions of two- and three-dimensional granules in bounded and
  unbounded domains?
\end{quotation}
In experiments, it is known that interactions between bounding walls and
granules cannot be ignored~\cite{qua-gan-you2021,
KaouiHartingMisbah2011_PRE, her-sto-gra2005}. Therefore, the necessary
modification to the code will be implemented so that confinement can be
considered. This modification requires introducing an additional layer
potential on the bounding walls, and elements of this modification have
already been incorporated (Figure~\ref{fig:stenosis}). However, to
consider rheological properties of the suspensions, a periodic domain,
such as a Couette device, will be considered. Such multiply-connected
domains require additional modifications to the IE formulation. PI
Quaife is qualified to perform this task since he has simulated
suspensions with confinement for several projects including dynamic
suspensions~\cite{qua-bir2014, bys-sha-qua2020} and porous
media~\cite{chi-moo-qua2020, dea-qua-bir-jua2018, moo-che-chi-qua2022,
qua-gan-you2021}.

@@

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 5.} How can IEs be coupled with a Newton method to
  solve the non-linear PDE~\eqref{eqn:phase}?
\end{quotation}
For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. This requires solving the variable coefficient PDE
$-\epsilon^2 \Delta v + W''(\phi^{N}) v = F[\phi^N]$, which is
challenging with an integral equation since a fundamental solution is
not readily available. However, by choosing a double-well potential with
$W''(\phi_0) = W''(\phi_1) = \beta > 0$, we propose to instead use an
inexact Newton's method where $(F')^{-1}$ is approximated with the
solution of $-\epsilon^2 \Delta v + \beta v = F[\phi^N]$. We test the
efficacy of the exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1) \subset \RR$ by discretizing $F'$
with a second-order finite difference stencil. The double-well potential
is illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the same solution. Figure~\ref{fig:CA}(b)
shows the norm of the residual at each Newton
\begin{wrapfigure}[18]{l}{1.8in}
  \vspace{-7pt}
  \centering
  \input{figures/doublewell.tikz} 
  \vspace{-10pt}
  \caption{\label{fig:CA} \footnotesize (a) A double-well potential. (b)
  The convergence of exact and inexact Newton's methods applied
  to~\eqref{eqn:phaseRoot}.}
\end{wrapfigure}
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.
